Technical Report

### 2017-09-29 esop-2018/paper.tex

\begin{figure}[t]
  \centering
  \subfloat{\usebox{\testit}}
  \subfloat{\usebox{\progfthree}}
  \caption{test}
\end{figure}
\end{comment}
\begin{figure}[t]
  \centering
  \begin{minipage}[t]{.37\linewidth}
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     ${\text{\tiny\itshape\text{(continues in next column)}}}$
\end{lstlisting}
  \end{minipage}%
  \begin{minipage}[t]{.37\linewidth}
\begin{lstlisting}
     ${\text{\tiny\itshape\text{(continued from previous column)}}}$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \end{minipage}%
  \begin{minipage}[t]{.25\linewidth}
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \end{minipage}
  \caption{ Two programs that rely on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to \texttt{f4}.}
  \label{fig:prog-f3-and-g1}
\end{figure}

>>>>

\begin{figure}[t]
  \centering
  \subfloat{\usebox{\testit}}
  \subfloat{\usebox{\progfthree}}
  \caption{test}
\end{figure}
\end{comment}
\begin{figure}[t]
  \centering
  \begin{minipage}[t]{.37\linewidth}
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     ${\text{\tiny\itshape(continues in next column)}}$
\end{lstlisting}
  \end{minipage}%
  \begin{minipage}[t]{.37\linewidth}
\begin{lstlisting}
     ${\text{\tiny\itshape(continued from previous column)}}$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \end{minipage}%
  \begin{minipage}[t]{.25\linewidth}
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \end{minipage}
  \caption{ Two programs that rely on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to \texttt{f4}.}
  \label{fig:prog-f3-and-g1}
\end{figure}

### 2017-09-29 fully-abstract-comp/technical-report.tex

\noindent\textbf{xjmp}\\
\lau{22-09-2017: Can we allow the callee to return any stack pointer? At the moment, it is not possible to create an empty stack pointer which may be a saving grace. If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
\begin{align*}
  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
                                   \begin{cases}
                                     \arraycolsep=0pt
                                     \begin{array}[t]{rl}
                                       \Phi & \updReg{r_1}{w_1}\\
                                            & \updReg{r_2}{w_2}\\
                                            & \updReg{\pcreg}{c_\code}\\
                                            & \updReg{\rdata}{c_\data}
                                     \end{array} &
                                     \begin{array}[t]{l}
                                       \Phi(r_1) = \sealed{c_\code,\sigma} \tand \\
                                       \Phi(r_2) = \sealed{c_\data,\sigma} \tand \\
                                       \nonExec{c_\data} \tand \\
                                       w_1 = \linCons{\Phi(r_1)} \tand \\
                                       w_2 = \linCons{\Phi(r_2)}
                                     \end{array}
                                     \\&\\
                                     \sourcecolor\left.
                                       \arraycolsep=0pt
                                       \begin{array}[t]{rl}
                                         \Phi' & \updReg{r_2}{0} \\
                                               & \updReg{\pcreg}{\opc}\\
                                               & \updReg{\rdata}{0}\\
                                               & \updReg{\rstk}{c_\stk}
                                       \end{array}\right.& 
                                     \sourcecolor\left.
                                       \begin{array}[t]{l}
                                         \Phi(r_1) = \retptrc(\sigma) \tand \\
                                         \Phi(r_2) = \retptrd(\sigma) \tand \\
                                         \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
                                         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
                                         \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
                                         \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
                                         c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
                                         \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
                                       \end{array}\right.
                                     \\
                                     \failed & \totherwise
                                   \end{cases}
\end{align*}

@@

One way to solve this problem is to make the "top addess" of the stack known. There are many ways to do this, but we have chosen to do the following:
\begin{itemize}
\item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
\item The base address of the stack capability is a fixed address, so in the semantice, it will be expressed as a constant that is publicly known.
\end{itemize}

>>>>

\noindent\textbf{xjmp}\\
\lau{22-09-2017: Can we allow the callee to return any stack pointer? At the moment, it is not possible to create an empty stack pointer which may be a saving grace. If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
\lau{28-09-2017: For now, the jump to a return pointer does the stack end point check. We could possibly move this to the proposed semantic condition as well.}
\begin{align*}
  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
                                   \begin{cases}
                                     \arraycolsep=0pt
                                     \begin{array}[t]{rl}
                                       \Phi & \updReg{r_1}{w_1}\\
                                            & \updReg{r_2}{w_2}\\
                                            & \updReg{\pcreg}{c_\code}\\
                                            & \updReg{\rdata}{c_\data}
                                     \end{array} &
                                     \begin{array}[t]{l}
                                       \Phi(r_1) = \sealed{c_\code,\sigma} \tand \\
                                       \Phi(r_2) = \sealed{c_\data,\sigma} \tand \\
                                       \nonExec{c_\data} \tand \\
                                       w_1 = \linCons{\Phi(r_1)} \tand \\
                                       w_2 = \linCons{\Phi(r_2)}
                                     \end{array}
                                     \\&\\
                                     \sourcecolor\left.
                                       \arraycolsep=0pt
                                       \begin{array}[t]{rl}
                                         \Phi' & \updReg{r_2}{0} \\
                                               & \updReg{\pcreg}{\opc}\\
                                               & \updReg{\rdata}{0}\\
                                               & \updReg{\rstk}{c_\stk}
                                       \end{array}\right.&
                                     \sourcecolor\left.
                                       \begin{array}[t]{l}
                                         \Phi(r_1) = \retptrc(\sigma) \tand \\
                                         \Phi(r_2) = \retptrd(\sigma) \tand \\
                                         \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
                                         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
                                         \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
                                         \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
                                         c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
                                         \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
                                       \end{array}\right.
                                     \\
                                     \failed & \totherwise
                                   \end{cases}
\end{align*}

@@

One way to solve this problem is to make the "top addess" of the stack known. There are many ways to do this, but we have chosen to do the following:
\begin{itemize}
\item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
\item The base address of the stack capability is a fixed address, so in the semantics, it will be expressed as a constant that is publicly known.
\item At some point before a call, it must be checked whether the stack we are using actually has the globally known base address (if not we must fail because we cannot trust this stack).
\item To ensure that the check is made, we require it in the semantic condition (at this moment of time not defined, so we are yet to see what it looks like).
\end{itemize}
